---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  warning=FALSE,
  message=FALSE
)
```

# time.table

<!-- badges: start -->
<!-- badges: end -->

Often when dealing with messy, real-world time series data, we have many variables sampled at different frequencies or sampled intermittently. The 'time.table' library provides functions to easily wrangle these different variables into a uniformly-spaced tabular format. 'time.table' is a library built around the popular high-performance data manipulation library 'data.table'.

## Installation

You can install the development version of time.table from [GitHub](https://github.com/bmgaldo/time.table) with:

``` r
# install.packages("devtools")
devtools::install_github("bmgaldo/time.table")
```

# Example Data 

```{r}
library(time.table)
library(data.table)
library(table.express)


df_admit_discharge = data.frame(patient_id=paste0("pat_",1:2),
                                admit=c(as.POSIXct("2012-03-12 12:12:12"),
                                        as.POSIXct("2012-03-13 11:11:11")),
                                discharge=c(as.POSIXct("2012-3-20 12:12:12"),
                                            as.POSIXct("2012-03-20 10:10:10")))

head(df_admit_discharge)
```

# Create TimeTable Template 

this specifics the start & end times of each IDs table. This template will be the scaffold in which variables will be added to a timetable.

```{r}

TimeTableTempl = time.table::GetTimeTableTemplate(start_times = df_admit_discharge$admit,
                                                  end_times = df_admit_discharge$discharge,
                                                  IDs = df_admit_discharge$patient_id,
                                                  sample_interval= "60 min" # grid point at every 60th minute 
)

head(TimeTableTempl)
```



# Add Static feature to a TimeTable

First, we'll generate some fake data for our two patients. In this example, a patient's heart rate is measured every few hours.

```{r}

# patient 1
times_pat_1 = seq(as.POSIXct("2012-03-12 12:12:12"),
                  as.POSIXct("2012-03-20 20:20:20"),
                  by="5 hours") # patient 1 has their heart rate measured every 5 hours
values_pat_1 = rnorm(length(times_pat_1),80,20)
ID_pat_1 = rep("pat_1",length(times_pat_1))

# patient 2
times_pat_2 = seq(as.POSIXct("2012-03-13 11:11:11"),
                  as.POSIXct("2012-03-20 10:10:10"),
                  by="4 hours") # patient 2 has their heart rate measured every 4 hours
values_pat_2 = rnorm(length(times_pat_2),70,10)
ID_pat_2 = rep("pat_2",length(times_pat_2))


# let's store consolidate these into a data.frame
df_heartrate = data.frame(IDs = c(ID_pat_1,ID_pat_2),
                          heart_rate = c(values_pat_1,values_pat_2),
                          times = c(times_pat_1,times_pat_2))

head(df_heartrate)

```

```{r}

heart_rate_tt = time.table::PutOnTimeTable(value = df_heartrate$heart_rate,
                                           datetimes = df_heartrate$times,
                                           IDs = df_heartrate$IDs,
                                           var_name = "heart_rate",
                                           Template = TimeTableTempl, # pass template from section one
                                           sample_interval = "60 min",
                                           aggregration_type = "mean", #how to handle multiple measurements for same time point
                                           duration  = "2 hours"# assume measures are carried forward 2 hours if no new measurement
)

head(heart_rate_tt)
```

# Add Static feature to a TimeTable

```{r}

# patient 1
times_pat_1 = as.POSIXct("2012-03-12 12:12:12")
values_pat_1 = "male"
ID_pat_1 = rep("pat_1",length(values_pat_1))

# patient 2
times_pat_2 = as.POSIXct("2012-03-13 11:11:11")
values_pat_2 = "female"
ID_pat_2 = rep("pat_2",length(values_pat_2))


# let's store consolidate these into a data.frame
df_sex = data.frame(IDs=c(ID_pat_1,ID_pat_2),
                    value=c(values_pat_1,values_pat_2),
                    datetimes=c(times_pat_1,times_pat_2))

head(df_sex)

sex_tt = PutOnTimeTable(value = df_sex$value,
                        datetimes = df_sex$datetimes,
                        IDs = df_sex$IDs,
                        var_name = "sex",
                        Template = TimeTableTempl, # pass template from section one 
                        sample_interval = "60 min", 
                        aggregration_type = "latest", # how to handle multiple measurements for same time point 
                        duration = NA # a measure doesn't expire here since it's static 
                        # both 'NA' or 'Inf' would work here.
)

head(sex_tt)
```

# Combine TimeTables into a data.table  

## Cbind
```{r}
# 
tt = CombineTimeTables(c("sex_tt","heart_rate_tt"),
                       join_type='cbind')

head(tt)
```

## Right Join
```{r}

CombineTimeTables(c("sex_tt","heart_rate_tt"),
                  join_type='right_join') 

```
